Cover Letter: Cudele (Paper 306)

### 2018-01-09 paper/sections/5_implementation.tex
paper: typos and revision 2
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

We use a programmable storage approach~\cite{sevilla:eurosys17} to design
Cudele; namely, we try to leverage components inside CephFS to inherit the
robustness and correctness of the internal subsystems. Using this `dirty-slate'
approach, we only had to implement 4 of the 6 mechanisms from
Figure~\ref{fig:decouple} and just 1 required changes to the underlying storage
system itself.  In this section, we first describe a CephFS internal subsystem or
component and then we show how we use it in Cudele.

@@

In CephFS, the metadata store is a data structure that represents the file
system namespace. This data structure is stored in two places: in memory ({\it
i.e.} in the collective memory of the metadata server cluster) and as objects
in the object store. In the object store, directories and their inodes are
stored together in objects to improve the performance of scans.  The metadata
store data structure is structured as a tree of directory fragments making it
easier to read and traverse. \textbf{In Cudele} the RPCs mechanism uses the
metadata store to service requests. Using code designed for recovery,
Volatile Apply and Nonvolatile Apply replay updates onto the metadata store in
memory and in the object store, respectively.  When the clients are ready to
merge their updates back into the global namespace, they pass a binary file of
metadata updates to the metadata server.

@@

The journal is the second way that CephFS represents the file system namespace;
it is a log of metadata updates that can materialize the namespace when the
updates are replayed onto the metadata store. The journal is a ``pile system";
writes are fast but reads are slow because state must be reconstructed.
Specifically, reads are slow because there is more state to read, it is
unorganized, and many of the updates may be redundant.  \textbf{In Cudele} the
journal format is used by Stream, Append Client Journal, Local Persist, and
Global Persist.  Stream is the default implementation for achieving global
durability in CephFS but the rest of the mechanisms are implemented by writing
with the journal format.  By writing with the same format, the metadata servers
can read and use the recovery code to materialize the updates from a client's
decoupled namespace ({\it i.e.} merge).  These implementations required no
changes to CephFS because the metadata servers know how to read the events the
library is writing.  By re-using the journal subsystem to implement the
namespace decoupling, Cudele leverages the write/read optimized data
structures, the formats for persisting events (similar to TableFS's
SSTables~\cite{ren:atc2013-tablefs}), and the functions for replaying events
onto the internal namespace data structures.

@@

The journal tool is used for disaster recovery and lets administrators view and
modify the journal. It can read the journal, export the journal as a file,
erase events, and apply updates to the metadata store.  To apply journal
updates to the metadata store, the journal tool reads the journal from object
store objects and replays the updates on the metadata store in the object
store.  \textbf{In Cudele} the external library the clients link into is based
on the journal tool. It already had functions for importing, exporting, and
modifying the updates in the journal so we re-purposed that code to implement
Append Client Journal, Volatile Apply, and Nonvolatile Apply.

@@

In CephFS, the inode cache reduces the number of RPCs between clients and
metadata servers. Without contention clients can resolve metadata reads locally
thus reducing the number of operations ({\it e.g.}, \texttt{lookup()}s).  For
example, if a client or metadata server is not caching the directory inode, all
creates within that directory will result in a lookup and a create request. If
the directory inode is cached then only the create needs to be sent.  The size
of the inode cache is configurable so as not to saturate the memory on the
metadata server -- inodes in CephFS are about 1400
bytes~\cite{docs:cephinternals}.  The inode cache has code for manipulating
inode numbers, such as pre-allocating inodes to clients. \textbf{ In
Cudele}, Nonvolatile Apply uses the internal inode cache code to allocate inodes
to clients that decouple parts of the namespace and to skip inodes used by the
client at merge time.

@@

In CephFS, inodes already store policies, like how the file is striped across
the object store or for managing subtrees for load balancing. These policies
adhere to logical partitionings of metadata or data, like Ceph pools and file
system namespace subtrees. To implement this, the namespace data structure has
the ability to recursively apply policies to subtrees and to isolate subtrees
from each other. \textbf{In Cudele} the large inodes also store consistency and
durability policies. This approach uses the File Type
interface from the Malacology programmable storage
system~\cite{sevilla:eurosys17} and it tells clients how to access the
underlying metadata. The underlying implementation stores executable code in
the inode that calls the different Cudele mechanisms. Of course, there are many
security and access control aspects of this approach but that is beyond the
scope of this paper.

>>>>

We use a programmable storage approach~\cite{sevilla:eurosys17} to design
Cudele; namely, we try to leverage components inside CephFS to inherit the
robustness and correctness of the internal subsystems. Using this ``dirty-slate"
approach, we only had to implement 4 of the 6 mechanisms from
Figure~\ref{fig:decouple} and just 1 required changes to the underlying storage
system itself.  In this section, we first describe a CephFS internal subsystem or
component and then we show how we use it in Cudele.

@@

In CephFS, the metadata store is a data structure that represents the file
system namespace. This data structure is stored in two places: in memory ({\it
i.e.} in the collective memory of the metadata server cluster) and as objects
in the object store. In the object store, directories and their inodes are
stored together in objects to improve the performance of scans.  The metadata
store data structure is structured as a tree of directory fragments making it
easier to read and traverse. \textbf{In Cudele,} the RPCs mechanism uses the
in-memory metadata store to service requests. Using code designed for recovery,
Volatile Apply and Nonvolatile Apply replay updates onto the metadata store in
memory and in the object store, respectively.

%When the clients are ready to
%merge their updates back into the global namespace, they pass a binary file of
%metadata updates to the metadata server.

@@

The journal is the second way that CephFS represents the file system namespace;
it is a log of metadata updates that can materialize the namespace when the
updates are replayed onto the metadata store. The journal is a ``pile system";
writes are fast but reads are slow because state must be reconstructed.
Specifically, reads are slow because there is more state to read, it is
unorganized, and many of the updates may be redundant.  \textbf{In Cudele,} the
journal format is used by Stream, Append Client Journal, Local Persist, and
Global Persist.  Stream is the default implementation for achieving global
durability in CephFS but the rest of the mechanisms are implemented by writing
with the journal format.  By writing with the same format, the metadata servers
can read and use the recovery code to materialize the updates from a client's
decoupled namespace ({\it i.e.} merge).  These implementations required no
changes to CephFS because the metadata servers know how to read the events the
library is writing.  By re-using the journal subsystem to implement the
namespace decoupling, Cudele leverages the write/read optimized data
structures, the formats for persisting events (similar to TableFS's
SSTables~\cite{ren:atc2013-tablefs}), and the functions for replaying events
onto the internal namespace data structures.

@@

The journal tool is used for disaster recovery and lets administrators view and
modify the journal. It can read the journal, export the journal as a file,
erase events, and apply updates to the metadata store.  To apply journal
updates to the metadata store, the journal tool reads the journal from object
store objects and replays the updates on the metadata store in the object
store.  \textbf{In Cudele,} the external library the clients link into is based
on the journal tool. It already had functions for importing, exporting, and
modifying the updates in the journal so we re-purposed that code to implement
Append Client Journal, Volatile Apply, and Nonvolatile Apply.

@@

In CephFS, the inode cache reduces the number of RPCs between clients and
metadata servers. Without contention clients can resolve metadata reads locally
thus reducing the number of operations ({\it e.g.}, \texttt{lookup()}s).  For
example, if a client or metadata server is not caching the directory inode, all
creates within that directory will result in a lookup and a create request. If
the directory inode is cached then only the create needs to be sent.  The size
of the inode cache is configurable so as not to saturate the memory on the
metadata server -- inodes in CephFS are about 1400
bytes~\cite{docs:cephinternals}.  The inode cache has code for manipulating
inode numbers, such as pre-allocating inodes to clients. \textbf{ In
Cudele,} Nonvolatile Apply uses the internal inode cache code to allocate inodes
to clients that decouple parts of the namespace and to skip inodes used by the
client at merge time.

@@

In CephFS, inodes already store policies, like how the file is striped across
the object store or for managing subtrees for load balancing. These policies
adhere to logical partitionings of metadata or data, like Ceph pools and file
system namespace subtrees. To implement this, the namespace data structure has
the ability to recursively apply policies to subtrees and to isolate subtrees
from each other. \textbf{In Cudele,} the large inodes also store consistency and
durability policies. This approach uses the File Type
interface from the Malacology programmable storage
system~\cite{sevilla:eurosys17} and it tells clients how to access the
underlying metadata. The underlying implementation stores executable code in
the inode that calls the different Cudele mechanisms. Of course, there are many
security and access control aspects of this approach but that is beyond the
scope of this paper.

### 2018-01-09 paper/sections/6_results.tex
paper: typos and revision 2
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\newcomment{\indent We graph standard deviations for three runs (sometimes
error bars are too small to see) and normalize results to make our results more
generally applicable to different hardware. We use a CloudLab cluster of 34
nodes connected with 10Gbit ethernet, each with 16 2.4 GHz CPUs, 128GB RAM, and
400GB SSDs.  Each node uses Ubuntu 14.04 and we develop on Ceph's Jewel
release, version 10.2.1, which was released in May 2016. We use 1 monitor
daemon, 3 object storage daemons, 1 metadata server daemon, and up to 20
clients.}  We scope the evaluation to one metadata server and scale the number
of parallel clients \newcomment{each doing 100K operations because 100K is the
maximum recommended size of a directory in CephFS. We scale to 20 clients
because, as shown in Section~\S\ref{sec:posix-overheads}, 20 clients is enough
to saturate the resources of a single metadata server.} This type of analysis
shows the capacity and performance of a metadata server with superior metadata
protocols, which should be used to inform \oldcomment{load
balancing}\newcomment{metadata distribution} across a \oldcomment{metadata}
cluster. Load balancing across a cluster of metadata servers with partitioning
and replication can be explored with something like
Mantle~\cite{sevilla:sc15-mantle}.  \newcomment{To make our results
reproducible, this paper adheres to The Popper
Convention~\cite{jimenez_popper_2016}, so experiments can be examined in more
detail, or even re-run, by visiting the \texttt{[source]} link next to each
figure.  The source code is available on the Cudele branch~\cite{docs:cudele}
of our Ceph fork.}

@@

\oldcomment{Figure~\ref{fig:composable-mechanisms} shows the runtime of the
Cudele mechanisms for a single client creating files in the same directory, }
\newcomment{We measure the overhead of each Cudele mechanism by having 1 client create
100K files in a directory for various subtree configurations.
Figure~\ref{fig:composable-mechanisms} shows the time that it takes each Cudele
mechanism to process all metadata events.  Results are} normalized to the time
it takes to write \oldcomment{100K file create} updates to the client's in-memory journal
({\it i.e.} the Append Client Journal mechanism), which is about 11K creates/sec.  \newcomment{The first graph
groups the consistency mechanisms, the second groups the durability mechanisms,
and the third has compositions representing real-world systems.} \oldcomment{Stream is an
approximation of the overhead and is calculated by subtracting the runtime of
the job with the journal turned off from the runtime with the journal turned
on.  100K is the maximum recommended size of a directory in CephFS;
preliminary experiments with larger directory sizes show memory problems.}

@@

\newcomment{\textbf{Cudele setup}: enable global durability with Stream and
strong consistency with RPCs to mirror the setup from the problem presented in
Figure~\ref{fig:overhead-b}. We configure one subtree with an interfere policy
of ``allow" and another subtree with ``block" so \texttt{-EBUSY} is
returned to interfering clients.  The former is the default behavior in file
systems and the latter isolates performance from malicious clients.}

\oldcomment{We only scale to 15 clients because the performance variability of
a nearly overloaded metadata server, in our case 18 clients, is too high.}
Figure~\ref{fig:block-allow} plots the overhead of the slowest
client\oldcomment{and the error bars are the standard deviation of 3 runs.
``Interfere (Fig 3c)" is the result from Figure~\ref{fig:overhead-b};
``Interfere" uses the allow/block API to return \texttt{-EBUSY} to interfering
clients; and ``Isolated" is the baseline performance without an interfering.
Each subtree has RPCs and Stream enabled.  Results are normalized to the
runtime of a single client creating files in a directory.  Note that IndexFS
has a similar behavior to ``interfere" with leases, except clients block.
}\newcomment{, normalized to 1 client that creates 100K files in isolation
(about 513 creates/sec). ``Interference" and ``no interference" is the
performance with and without an interfering client touching files in every
directory, respectively. The goal is to explicitly isolate clients so that
performance is similar to the ``no interference" curve, which has lower
slowdowns (on average, \(1.42\times\) per client compared to \(1.67\times\) per
client for ``interfere") and less variability (on average, a standard deviation
of 0.06 compared to 0.44 for ``interfere").  ``Block interference" uses the
Cudele API to block interfering clients and the slowdown (\(1.34\times\) per
client) and variability (0.09) look very similar to ``no interference" for a
larger number of clients. For smaller clusters the overhead to reject requests
is more evident when the metadata server is underloaded so the slowdowns are
similar to ``interference".  We conclude that administrators can block interfering clients to get the same
performance as isolated scenarios but there is a non-negligible overhead for
rejecting requests when the metadata server is not operating at peak
efficiency.}

@@

The final use case shows how the API gives
\oldcomment{users}\newcomment{administrators} fine-grained control of the
consistency semantics \newcomment{to support current practices and scientific
workflows in HPC}. \oldcomment{The use case is that u}\newcomment{U}sers often
leverage the file system to check the progress of jobs using \texttt{ls} even
though this operation is notoriously heavy-weight~\cite{carns:ipdps09-pvfs,
eshel:fast10-panache}. The number of files or size of the files is indicative
of the progress. This practice is not too different from cloud systems that use the
file system to manage the progress of jobs; Spark/Hadoop writes to temporary
files, renames them when complete, and creates a ``DONE" file to indicate to
the runtime that the task did not fail and should not be re-scheduled on
another node. \newcomment{Also, the browser interface lets Hadoop/Spark users check
progress by querying the file system and returning a \% of job complete
metric.}

>>>>

\newcomment{\indent We graph standard deviations for three runs (sometimes
error bars are too small to see) and normalize results to make our results more
generally applicable to different hardware. We use a CloudLab cluster of 34
nodes connected with 10Gbit ethernet, each with 16 2.4 GHz CPUs, 128GB RAM, and
400GB SSDs.  Each node uses Ubuntu 14.04 and we develop on Ceph's Jewel
release, version 10.2.1, which was released in May 2016. We use 1 monitor
daemon, 3 object storage daemons, 1 metadata server daemon, and up to 20
clients.}  We scope the evaluation to one metadata server and scale the number
of parallel clients \newcomment{each doing 100K operations because 100K is the
maximum recommended size of a directory in CephFS. We scale to 20 clients
because, as shown in Section~\S\ref{sec:posix-overheads}, 20 clients is enough
to saturate the resources of a single metadata server.} This type of analysis
shows the capacity and performance of a metadata server with superior metadata
protocols, which should be used to inform \oldcomment{load
balancing}\newcomment{metadata distribution} across a \oldcomment{metadata}
cluster. Load balancing across a cluster of metadata servers with partitioning
and replication can be explored with something like
Mantle~\cite{sevilla:sc15-mantle}.  \newcomment{To make our results
reproducible, this paper adheres to The Popper
Convention~\cite{jimenez_popper_2016} so experiments can be examined in more
detail, or even re-run, by visiting the \texttt{[source]} link next to each
figure.  The source code for Cudele is available on a branch~\cite{docs:cudele}
of our Ceph fork.}

@@

\oldcomment{Figure~\ref{fig:composable-mechanisms} shows the runtime of the
Cudele mechanisms for a single client creating files in the same directory, }
\newcomment{We measure the overhead of each Cudele mechanism by having 1 client create
100K files in a directory for various subtree configurations.
Figure~\ref{fig:composable-mechanisms} shows the time that it takes each Cudele
mechanism to process all metadata events.  Results are} normalized to the time
it takes to write \oldcomment{100K file create}updates to the client's in-memory journal
({\it i.e.} the Append Client Journal mechanism), which is about 11K creates/sec.  \newcomment{The first graph
groups the consistency mechanisms, the second groups the durability mechanisms,
and the third has compositions representing real-world systems.} \oldcomment{Stream is an
approximation of the overhead and is calculated by subtracting the runtime of
the job with the journal turned off from the runtime with the journal turned
on.  100K is the maximum recommended size of a directory in CephFS;
preliminary experiments with larger directory sizes show memory problems.}

@@

\newcomment{\textbf{Cudele setup}: enable global durability with Stream and
strong consistency with RPCs to mirror the setup from the problem presented in
Figure~\ref{fig:overhead-b}. We configure one subtree with an interfere policy
of ``allow" and another subtree with ``block" so \texttt{-EBUSY} is
returned to interfering clients.  The former is the default behavior in file
systems and the latter isolates performance from interfering clients.}

\oldcomment{We only scale to 15 clients because the performance variability of
a nearly overloaded metadata server, in our case 18 clients, is too high.}
Figure~\ref{fig:block-allow} plots the overhead of the slowest
client\oldcomment{and the error bars are the standard deviation of 3 runs.
``Interfere (Fig 3c)" is the result from Figure~\ref{fig:overhead-b};
``Interfere" uses the allow/block API to return \texttt{-EBUSY} to interfering
clients; and ``Isolated" is the baseline performance without an interfering.
Each subtree has RPCs and Stream enabled.  Results are normalized to the
runtime of a single client creating files in a directory.  Note that IndexFS
has a similar behavior to ``interfere" with leases, except clients block.
}\newcomment{, normalized to 1 client that creates 100K files in isolation
(about 513 creates/sec). ``Interference" and ``no interference" is the
performance with and without an interfering client touching files in every
directory, respectively. The goal is to explicitly isolate clients so that
performance is similar to the ``no interference" curve, which has lower
slowdowns (on average, \(1.42\times\) per client compared to \(1.67\times\) per
client for ``interference") and less variability (on average, a standard deviation
of 0.06 compared to 0.44 for ``interference").  ``Block interference" uses the
Cudele API to block interfering clients and the slowdown (\(1.34\times\) per
client) and variability (0.09) look very similar to ``no interference" for a
larger number of clients. For smaller clusters the overhead to reject requests
is more evident when the metadata server is underloaded so the slowdowns are
similar to ``interference".  We conclude that administrators can block interfering clients to get the same
performance as isolated scenarios but there is a non-negligible overhead for
rejecting requests when the metadata server is not operating at peak
efficiency.}

@@

The final use case shows how the API gives
\oldcomment{users}\newcomment{administrators} fine-grained control of the
consistency semantics \newcomment{to support current practices and scientific
workflows in HPC}. \oldcomment{The use case is that u}\newcomment{U}sers often
leverage the file system to check the progress of jobs using \texttt{ls} even
though this operation is notoriously heavy-weight~\cite{carns:ipdps09-pvfs,
eshel:fast10-panache}. The number of files or size of the files is indicative
of the progress. This practice is not too different from cloud systems that use the
file system to manage the progress of jobs; Spark/Hadoop writes to temporary
files, renames them when complete, and creates a ``DONE" file to indicate to
the runtime that the task did not fail and should not be re-scheduled on
another node. \newcomment{So the browser interface lets Hadoop/Spark users check
progress by querying the file system and returning a \% of job complete
metric.}

### 2018-01-09 paper/sections/7_conclusion.tex
paper: typos and revision 2
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

%This approach is superior to mounting different storage systems in a global
%namespace because there is no need to provision dedicated storage clusters to
%applications or move data between these systems.  
Cudele prompts many avenues for future work.  First is to co-locate HPC
workflows with real highly parallel runtimes from the cloud in the same
namespace. This setup shows how Cudele reasonably incorporates both programming
models (client-driven parallelism and user-defined workflows) at the same time
and should show large performance gains.  Second is dynamically changing
semantics of a subtree from stronger to weaker guarantees (or vice versa). This
reduces data movement across storage cluster and file system boundaries so
the results of a Hadoop job do not need to be migrated into CephFS for
other processing; instead the \oldcomment{user}\newcomment{administrator} can
change the semantics of the HDFS subtree into a CephFS subtree, which may cause
metadata/data movement to ensure strong consistency. Third is embeddable
policies, where child subtrees have specialized features but still maintain
guarantees of their parent subtrees.  For example, a RAMDisk subtree is POSIX
IO-compliant but relaxes durability constraints, so it can reside under a POSIX
IO subtree alongside a globally durable subtree.

>>>>

%This approach is superior to mounting different storage systems in a global
%namespace because there is no need to provision dedicated storage clusters to
%applications or move data between these systems.  
Cudele prompts many avenues for future work.  First is to co-locate HPC
workflows with real highly parallel runtimes from the cloud in the same
namespace. This setup would show how Cudele reasonably incorporates both programming
models (client-driven parallelism and user-defined workflows) at the same time
and should show large performance gains.  Second is dynamically changing
semantics of a subtree from stronger to weaker guarantees (or vice versa). This
reduces data movement across storage cluster and file system boundaries so
the results of a Hadoop job do not need to be migrated into CephFS for
other processing; instead the \oldcomment{user}\newcomment{administrator} can
change the semantics of the HDFS subtree into a CephFS subtree, which may cause
metadata/data movement to ensure strong consistency. Third is embeddable
policies, where child subtrees have specialized features but still maintain
guarantees of their parent subtrees.  For example, a RAMDisk subtree is POSIX
IO-compliant but relaxes durability constraints, so it can reside under a POSIX
IO subtree alongside a globally durable subtree.
